---
title: "Final Project"
author: Tori Macfarlan
output: pdf_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(width = 100)
```

## Introduction

Do leaders really impact their governments that significantly? Does the death of a leader, especially in an assassination, frequently lead to a change in the type of government, or do the histories and traditions of a country play a larger role in its shaping? This project will seek to demonstrate a leader's assassination does not have a significant impact on the polity of a country following the event, but an assassination attempt (regardless of success) cause with an increase in civil wars and/or international conflicts. This discussion also theorizes less democratic governments cause an increase in turmoil following an assassination attempt.  

## Data
```{r}
leaders <- read.csv("data/leaders.csv")


leaders$result <- factor(leaders$result)

leaders$result.bin <- ifelse(as.numeric(leaders$result) == 1 | as.numeric(leaders$result) == 2 | as.numeric(leaders$result) == 3 | as.numeric(leaders$result) == 4, "TRUE" , "FALSE")


leaders$polity.change <- (leaders$polityafter - leaders$politybefore)

hist(leaders$polity.change , main = "Polity Change Following Assassination Attempts" , xlab = "Significance of Polity Change")

```
The data for this discussion comes from the Polity Project. This organization evaluated regimes from `r range(leaders$year)[1] ` to `r range(leaders$year)[2] ` and quantified the type of government "from -10 (hereditary monarchy) to 10 (consolidated democracy)" (Imai, Kosuke. QSS). The data set contains leaders and governments that experienced one or more assassination attempts. The data evaluates the type of regime prior to and after the assassination attempts. The data also notates whether or not a country was involved in civil war or international conflict before and after assassination attempts. This data set is particularly useful for evaluating the questions presented in this discussion because it uses expert coding to quantify the type of government, a categorical variable. The observations in the data set contain characteristics that are diverse and roughly comparable to the actual population (such as age of the leader, year of the assassination attempt) in order to account for potential confounding factors. 

The graph above depicts the extent to which the type of government changed following an assassination attempt. A cursory glance seems to support the hypothesis that the type of government did not change significantly following an assassination attempt, but we will evaluate this in greater detail below.



## Results

We will evaluate the change in regime type as a function of whether or not a government's leader died in an assassination attempt. Given the success or failure of an assassination attempt is considered random in this data set, the extent to which a leader's assassination results in a significant polity change will be evaluated. To support the hypothesis above, there should be a low correlation between assassination of a leader and the resulting polity change. 



```{r}
death_fit <- lm(leaders$polity.change ~ leaders$result)

death_fit_sum <- summary(death_fit)

plot(leaders$result, leaders$polity.change, main = "Polity Changes Following an Assassination Attempt" , xlab = "Assassination Results" , ylab = "Polity Change", xaxt= 'n', bty='L')
axis(1, seq(0,10,1))

knitr::kable(data.frame(`X-Axis Value`=c(1:10), Results=levels(leaders$result)), label="label2")
```
The chart above shows more significant policy changes in the section of the chart where the leader died (Assassination Results 1-4). However, there is still some polity change in the remainder of the graph where the leader survived.

```{r}
death_fit_bin <- lm(leaders$polity.change ~ leaders$result.bin)

death_fit_bin_sum <- summary(death_fit_bin)

boxplot(leaders$polity.change ~ leaders$result.bin, main = "Polity Changes Following an Assassination Attempt" , xlab = "Assassination Attempt Success" , ylab = "Polity Change")

```

```{r, results='asis', message = FALSE}
stargazer::stargazer(death_fit_bin, header = FALSE, label="label Binary", dep.var.caption = "Polity Change")

```



In addition to the change in polity, we are evaluating the extent to which the countries in the data experienced increased turmoil following an assassination attempt. If the secondary hypothesis above is correct, an increase in turmoil (civil war OR international war) will correlate with successful assassinations. 

```{r}

leaders$turmoil <- (leaders$interwarafter + leaders$civilwarafter) - (leaders$interwarbefore + leaders$civilwarbefore)
leaders$turmoil_before <- (leaders$interwarbefore + leaders$civilwarbefore)
leaders$turmoil_after <- (leaders$interwarafter + leaders$civilwarafter)

turmoil_fit <- lm(leaders$turmoil ~ leaders$result.bin)
turmoil_fit_sum <- summary(turmoil_fit)
plot(leaders$result, leaders$turmoil, main = "Turmoil Following an Assassination Attempt" , xlab = "Assassination Results" , ylab = "Turmoil Change", xaxt= 'n')
axis(1, seq(0,10,1))

knitr::kable(data.frame(`X-Axis Value`=c(1:10), Results=levels(leaders$result)), label="label1")

```
The data above shows a statistically significant correlation between the results of an assassination attempts and the change in turmoil after. Leader deaths that occur immediately following the assassination (within a day) and over a longer period (between a week an a month) correlated with a decrease in overall turmoil. This particular statistic appears counter-intuitive, particularly given we do not see the same trend in leaders who died between a week and a day. While a loose correlation exists, we should look for confounding variables in the data or find a way to control for potentially unknown confounding variables.

Below we look at different types of governments (prior to an assassination attempt) and evaluate the relationship between polity and turmoil. 

```{r}

turmoil_pretype <- lm(leaders$turmoil ~ leaders$politybefore)
turmoil_pretype_sum <- summary(turmoil_pretype)

plot(leaders$politybefore, leaders$turmoil, main = "Impact of Government Type on Turmoil" , xlab = "Type of Government" , ylab =  "Change in Turmoil")

```

```{r, results='asis', message = FALSE}
stargazer::stargazer(turmoil_pretype, header = FALSE, label="label3", dep.var.caption = "Change in Turmoil")

```
The graph and table above show a low correlation between the type of government prior to the assassination attempt and the resulting turmoil.


```{r}

turmoil_preturmoil <- lm(leaders$turmoil ~ leaders$turmoil_before )
turmoil_preturmoil_sum <- summary(turmoil_preturmoil)

plot(leaders$turmoil, leaders$turmoil_before, main = "Impact of Turmoil Before on Turmoil Change" , xlab = "Turmoil Before" , ylab =  "Change in Turmoil")

```

```{r, results='asis', message = FALSE}
stargazer::stargazer(turmoil_preturmoil, header = FALSE, label="label4", dep.var.caption = "Change in Turmoil")

```
Table 4 and the above chart demonstrate a higher correlation. Here the r-squared value is .264, which is a relatively high correlation for statistical analysis in the social sciences. Further, the p-value is less than 0.01, indicating there is a very low probability of this correlation appearing by chance. The low p-value allows us to reject the null hypothesis. Turmoil prior to an assassination attempt seems to have a strong causality to the turmoil following an assassination attempt, casting significant doubt on our secondary hypothesis (that assassination attempts cause turmoil).


```{r, message = FALSE}
library(tidyverse)
n_by_year <- leaders %>% group_by(year) %>% summarize(n = n())

plot(n_by_year$year, n_by_year$n)

n_by_year_lm <- lm(n_by_year$n ~ n_by_year$year)
n_by_year_lm_sum <- summary(n_by_year_lm)

```
```{r, results='asis', message = FALSE}
stargazer::stargazer(n_by_year_lm, header = FALSE, label="label5")

```
## Conclusion
The dependent variable data generally fits into a normal distribution, with the highest frequency of data changes occurring in the middle section, showing the least significant shift in polity following an assassination attempt. (Still working with file, more analysis to follow)


