---
title: "Final Project"
author: Tori Macfarlan
output: pdf_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(width = 100)
```

## Introduction

How much does a leader impact the policies of their country? Does the death of a leader, especially an assassination, frequently lead to a change in the type of government, or do the histories and traditions of a country play a larger role in its shaping? This project will demonstrate a leader's assassination does not have a significant impact on the resulting polity of a country, but an assassination attempt (regardless of success) correlates with an increase in civil wars and/or international conflicts. It is possible the type of government that makes assassination attempts more likely is also the type of government more inclined towards political turmoil in general (though an assassination attempt could exacerbate it). 

## Data
```{r}
leaders <- read.csv("data/leaders.csv")


leaders$result <- factor(leaders$result)

leaders$result.bin <- ifelse(as.numeric(leaders$result) == 1 | as.numeric(leaders$result) == 2 | as.numeric(leaders$result) == 3 | as.numeric(leaders$result) == 4, "TRUE" , "FALSE")


leaders$polity.change <- (leaders$polityafter - leaders$politybefore)

hist(leaders$polity.change , main = "Polity Change Following Assassination Attempts" , xlab = "Significance of Polity Change")

```
The data for this discussion comes from the Polity Project. This organization evaluated regimes from `r range(leaders$year)[1] ` to `r range(leaders$year)[2] ` and quantified the type of government "from -10 (hereditary monarchy) to 10 (consolidated democracy)" (Imai, Kosuke. QSS). The data set contains leaders who did not experience assassination attempts and leaders who experienced one or more assassination attempts. The data evaluates the type of regime prior to and after (if applicable) any assassination attempts. The data also notates whether or not a country was involved in civil war or international conflict before and after assassination attempts. 

The graph above depicts the extent to which the type of government changed following an assassination attempt. The data includes governments whose leaders did not experience an assassination attempt (*considering splitting the data set up in this section to show countries that did experience an assassination attempt juxtaposed to those that did not).


## Results

Below, we evaluate the change in regime type as 

a results section that contains a scatterplot of the main relationship of interest and output for the main regression of interest

```{r}




```


## Conclusion
The data generally fits into a normal distribution, with the highest frequency of data changes occurring in the middle section, showing the least significant shift in polity following an assassination attempt. 


