---
title: "Final Project"
author: Tori Macfarlan
output: pdf_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(width = 100)
```

## Introduction

How much does a leader impact the policies of their country? Does the death of a leader, especially an assassination, frequently lead to a change in the type of government, or do the histories and traditions of a country play a larger role in its shaping? This project will demonstrate a leader's assassination does not have a significant impact on the resulting polity of a country, but an assassination attempt (regardless of success) correlates with an increase in civil wars and/or international conflicts. It is possible the type of government that makes assassination attempts more likely is also the type of government more inclined towards political turmoil in general (though an assassination attempt could exacerbate it). 

## Data
```{r}
leaders <- read.csv("data/leaders.csv")
sum( is.na(leaders) )
sum( leaders == "" )
ls(leaders)
results <- factor(leaders$result)
levels(results)
#[1] "dies between a day and a week"              
# [2] "dies between a week and a month"            
 #[3] "dies within a day after the attack"         
# [4] "dies, timing unknown"                       
 #[5] "hospitalization but no permanent disability"
# [6] "not wounded"                                
 #[7] "plot stopped"                               
# [8] "survives but wounded severely"              
# [9] "survives, whether wounded unknown"          
#[10] "wounded lightly"  

leaders$result <- factor(leaders$result)

leaders$result.bin <- ifelse(as.numeric(leaders$result) == 1 | as.numeric(leaders$result) == 2 | as.numeric(leaders$result) == 3 | as.numeric(leaders$result) == 4, "TRUE" , "FALSE")

range(leaders$year)

```
The data for this discussion comes from the Polity Project. This organization evaluated regimes from `r range(leaders$year)[1] ` to `r range(leaders$year)[2] ` and quantified the type of government "from −10 (hereditary monarchy) to 10 (consolidated democracy)" (Imai, Kosuke. QSS). The data set contains leaders who did not experience assassination attempts and leaders who experienced one or more assassination attempts. The data evaluates the type of regime prior to and after (if applicable) any assassination attempts. The data also notates whether or not a country was involved in civil war or international conflict before and after assassination attempts. 




a data section that briefly describes the data source, describes how the key dependent and independent variables are measured (e.g., a survey, statistical model, or expert coding), and also produces a plot that summarizes the dependent variable

## Results

Below, we evaluate the change in regime type as 

a results section that contains a scatterplot of the main relationship of interest and output for the main regression of interest

## Conclusion

a brief (one paragraph) concluding section that summarizes your results, assesses the extent to which you find support for your hypothesis, describes limitations of your analysis and threats to inference, and states how your analysis could be improved (e.g., improved data that would be useful to collect)


